---
title: "DS 3606 Project 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
Youtubelink = "https://youtu.be/mD79gMNnW_E"
```

## Configure necessairy libraries
```{r}
library(MASS)
library(epitools)
library(naniar)
library(tidyverse)
library(mice)
library(ggplot2)
library(htmlwidgets)
library(caTools)
library(class)
library(caret)
library(GGally)
library(readr)
library(e1071)
library(naivebayes)
library(randomForest)
library(Metrics)
library(reshape)
library(knitr)
```

## Read in the Data
```{r}
employee_data = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv")
no_attrition_data = read.csv("https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2CompSet%20No%20Attrition.csv")
no_salary_data = read.csv("https://raw.githubusercontent.com/SucrazyGit/DS-3606-Project-2/main/CaseStudy2_NoSalary.csv")
```

## Summarise Data
```{r}
summary(employee_data) ## Summary statistics
gg_miss_var(employee_data) ## Looks for missing values
```


## Setting Variables to the correct format
```{r}
employee_train = employee_data[,-1]## Removes the ID term from the Data Set
employee_train = employee_train[,-22] ## Removes "Over 18" Variable
employee_train$BusinessTravel = as.factor(employee_train$BusinessTravel)
employee_train$Department = as.factor(employee_train$Department)
employee_train$JobLevel = as.factor(employee_train$JobLevel)
employee_train$EducationField = as.factor(employee_train$EducationField)
employee_train$Gender = as.factor(employee_train$Gender)
employee_train$JobRole = as.factor(employee_train$JobRole)
employee_train$MaritalStatus = as.factor(employee_train$MaritalStatus)
employee_train$OverTime = as.factor(employee_train$OverTime)
employee_train$Attrition = as.factor(employee_train$Attrition)
```

## Setting Variables to the correct format
```{r}
no_attrition_test = no_attrition_data[,-1]  ## Removes the ID term from the Data Set
no_attrition_test$BusinessTravel = as.factor(no_attrition_test$BusinessTravel)
no_attrition_test$Department = as.factor(no_attrition_test$Department)
no_attrition_test$JobLevel = as.factor(no_attrition_test$JobLevel)
no_attrition_test$EducationField = as.factor(no_attrition_test$EducationField)
no_attrition_test$Gender = as.factor(no_attrition_test$Gender)
no_attrition_test$JobRole = as.factor(no_attrition_test$JobRole)
no_attrition_test$MaritalStatus = as.factor(no_attrition_test$MaritalStatus)
no_attrition_test$OverTime = as.factor(no_attrition_test$OverTime)
```

## Setting Variables to the correct format
```{r}
no_salary_test = no_salary_data[,-1]  ## Removes the ID term from the Data Set
no_salary_test$BusinessTravel = as.factor(no_salary_test$BusinessTravel)
no_salary_test$Department = as.factor(no_salary_test$Department)
no_salary_test$EducationField = as.factor(no_salary_test$EducationField)
no_salary_test$Gender = as.factor(no_salary_test$Gender)
no_salary_test$JobLevel = as.factor(no_salary_test$JobLevel)
no_salary_test$JobRole = as.factor(no_salary_test$JobRole)
no_salary_test$MaritalStatus = as.factor(no_salary_test$MaritalStatus)
no_salary_test$OverTime = as.factor(no_salary_test$OverTime)
no_salary_test$Attrition = as.factor(no_salary_test$Attrition)
```

```{r}
corr_data = employee_train[, c(1,4,6,7.9,10,11,13,14,17,19,20,21,23,24,25,27,28,29,30,31,32,33,34)] ## Puts on the continuous variables into a data set
cormat = round(cor(corr_data),2) ## Setting up data set
melt_cormat = melt(cormat) ## Setting up data set
melt_cormat %>% ggplot(aes(x=X1, y=X2, fill=value)) + geom_tile() + ggtitle("Correlation Between Variables") ## Create correlation matrix
```

## Important Graphs
```{r}
##Graphs for important variables in relation to Attrition
employee_train %>% ggplot(aes(x=Age, color = Attrition)) + geom_boxplot() + ggtitle("Age vs Attrition") ## Plots age vs attrition
employee_train %>% ggplot(aes(x=MonthlyIncome, color = Attrition)) + geom_boxplot() + ggtitle("Monthly Income vs Attrition") ## Plots monthly income vs attrition
employee_train %>% ggplot(aes(x=TotalWorkingYears, color = Attrition)) + geom_boxplot() + ggtitle("Total Working Years vs Attrition") ## Plots total working years vs attrition
employee_train %>% ggplot(aes(x=YearsAtCompany, color = Attrition)) + geom_boxplot() + ggtitle("Years at the Company vs Attrition") ## Plots years at the company vs attrition
employee_train %>% ggplot(aes(x=DailyRate, color = Attrition)) + geom_boxplot() + ggtitle("Daily Rate vs Attrition") ## Plots daily rate vs attrition
employee_train %>% ggplot(aes(x=DistanceFromHome, color = Attrition)) + geom_boxplot() + ggtitle("Distance from Home vs Attrition") ## Plots distance from home vs attrition

##Graphs for important variables in relation to Monthly Income
employee_train %>% ggplot(aes(x=MonthlyIncome, color = JobLevel)) + geom_boxplot() + ggtitle("Salary vs Job Level") ## Plots monthly income vs job level
employee_train %>% ggplot(aes(x=MonthlyIncome, color = JobRole)) + geom_boxplot() + ggtitle("Salary vs Job Role") ## Plots monthly income vs job role
employee_train %>% ggplot(aes(x=MonthlyIncome, color = Gender)) + geom_boxplot() + ggtitle("Salary vs Gender") ## Plots monthly income vs job level
employee_train %>% ggplot(aes(x=MonthlyIncome, y=TotalWorkingYears, color=Gender)) + geom_point() + ggtitle("Salary vs Total Working Years") ## Plots monthly income vs total working years
employee_train %>% ggplot(aes(x=MonthlyIncome, color = BusinessTravel)) + geom_boxplot() + ggtitle("Salary vs Travel Amount") ## Plots monthly income vs travel amount
```

## Splits Data into test and training date sets
```{r}
set.seed(10000) ## Sets seed
splitPerc = .70 ## Determines test and train data split

trainIndices = sample(1:dim(employee_train)[1],round(splitPerc * dim(employee_train)[1])) ## Splits data set into training and test sets
train = employee_train[trainIndices,]
test = employee_train[-trainIndices,]
```

## Naive Bayes Model: Determines Attrition 
```{r}
nb_model = naive_bayes(as.factor(Attrition)~., data = employee_train) ## creates naive bayes model
nb_classifications = predict(nb_model,test) ## Predicts using model 
table(test$Attrition,nb_classifications) ## Generates the table of predicted vs actual
nb_CM = confusionMatrix(table(test$Attrition,nb_classifications)) ## Creates a confusion matrix
print(nb_CM) ## Prints all the models statistics including sensitivity and specificity 
```

## Random Forrest Model: Determines Attrition
```{r}
rf_model = randomForest(as.factor(Attrition)~., data = train) ## creates random forest model
rf_classifications = predict(rf_model,test) ## Predicts using the random forest model
table(test$Attrition,rf_classifications) ## Generates the table of predicted vs actual
rf_CM = confusionMatrix(table(test$Attrition,rf_classifications)) ## Creates a confusion matrix
print(rf_CM) ## Prints all the models statistics including sensitivity and specificity 
```

## Using Random Forest Model to predict attrition
```{r}
attrition_pred = predict(rf_model, no_attrition_test) ## Predicts the attrition classifications of the data set that doesn't have attrition values. Uses the random forest model to handle the predictions
final_attrition_data = cbind(no_attrition_data, attrition_pred) ## Combines predicted attrition values with the data set with no attrition to have a finalized data set
write.csv(final_attrition_data, "C:/Users/bmsuc/OneDrive/Desktop/Case2PredictionsBrandonSucreseAttrition.csv") ## Saves the new data set to the file location given
```

## Simple Linear Model: Determines Salary
```{r}
lm_model = lm(MonthlyIncome~.,data = train) ## creates simple linear regression model
coef(lm_model) ## Looks at the coefficients of the variables being utilized
summary(lm_model) ## Summaries the new linear model
lm_predictions = predict(lm_model,test) ## Predictions monthly income of the test set using the linear model
RMSE(lm_predictions, test$MonthlyIncome) ## Prints the RSME of the model
```

## Using Simple Linear Model to predict Monthly Income
```{r}
salary_pred = predict(lm_model, no_salary_test) ## Predicts the monthly income of the incomplete salary data set
final_salary_data = cbind(no_salary_data, salary_pred) ## Combines the incomplete salary data set and predicts into a single completed data set
write.csv(final_salary_data, "C:/Users/bmsuc/OneDrive/Desktop/Case2PredictionsBrandonSucreseSalary.csv") ## Saves the new data set to the file location given
```
